IMF and Benchmark Forecasts

### 2023-09-12 manuscript/manuscript.Rmd
manuscript daily update

---
title: "IMF and Benchmark Forecasts"
header-includes:
    - \usepackage{lineno}
    - \linenumbers
    -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
linestretch: 2 
bibliography:
  - references.bib
csl: vancouver.csl
---

@@

# Scores, by estimation method, Horizon and forecast source 
```{r, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

@@

#bring into long format
gather_scores <- all_scores |>
  data.table::copy() |>
  .d(target == "pcpi_pch") |>
  .d(,target := NULL) |>
  .d(, .(horizon, method, source, score, interval_score)) |>
  setnames("score", "sample_crps") |>
  melt(id.vars = c("horizon", "method", "source"), variable.name = "score_type", value.name = "score") |>
  .d(, combms := paste0(method,"_", score_type)) |>
  .d(, .(horizon, source, combms, score)) |>
  .d(, horizon := paste0("horizon = ", horizon)) |>
  .d(, score := round(score, 3)) |> #enough to distinguish
  dcast(horizon + combms ~ source, value.var = "score")

gather_scores |>
  data.table::copy() |>
  gt(rowname_col = "combms", groupname_col = "horizon") |>
  cols_label(
      IMF = "IMF",
      ar = "ar",
      bvar = "bvar"
    )

col_names <- c("IMF", "ar", "bvar")

@@

```{r, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

@@

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("interval", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score") |>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_is[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score)) +
    guides(fill = FALSE) +
    ggtitle(paste0("horizon = ", hor, " - Weighted Interval Score"))
}

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("crps", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score")|>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_cr[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score))+
    guides(fill = FALSE)+
    ggtitle(paste0("horizon = ", hor, " - CRPS by sample"))
}

@@

```{r cvgdat2, include=FALSE, echo = FALSE}

@@

```{r hey3, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

all_scores <- 
  crps_scores[wis_scores, on = c("source", "error_method", "method", "target", "horizon")]|>
  .d(error_method == "absolute") |>
  .d(, error_method := NULL)

gather_scores <- all_scores |>
  data.table::copy() |>
  .d(target == "ngdp_rpch") |>
  .d(,target := NULL) |>
  .d(, .(horizon, method, source, score, interval_score)) |>
  setnames("score", "sample_crps") |>
  melt(id.vars = c("horizon", "method", "source"), variable.name = "score_type", value.name = "score") |>
  .d(, combms := paste0(method,"_", score_type)) |>
  .d(, .(horizon, source, combms, score)) |>
  .d(, horizon := paste0("horizon = ", horizon)) |>
  .d(, score := round(score, 3)) |>
  dcast(horizon + combms ~ source, value.var = "score")

gather_scores |>
  data.table::copy() |>
  gt(rowname_col = "combms", groupname_col = "horizon") |>
  cols_label(
      IMF = "IMF",
      ar = "ar",
      bvar = "bvar"
    )

col_names <- c("IMF", "ar", "bvar")

```{r, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

@@

>>>>

---
title: "IMF and Benchmark Forecasts"
header-includes:
    - \usepackage{lineno}
    - \linenumbers
    -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
linestretch: 2 
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
bibliography:
  - references.bib
csl: vancouver.csl
---

@@

```{r loaddata, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

@@

# Scores, by estimation method, Horizon and forecast source 
```{r, cpitable = FALSE}
#bring into long format
long_scores <- gather_scores(all_scores, "pcpi_pch")

scores_to_table(all_scores, "pcpi_pch")

@@

```{r cpitileplot, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

@@

hor_list <- as.list(1:4)

plots_is <- lapply(hor_list, function(i) tile_plot(long_scores, hor = (i-1)*0.5, "interval"))
plots_cr <- lapply(hor_list, function(i) tile_plot(long_scores, hor = (i-1)*0.5, "crps"))

@@

```{r gdptable, echo = FALSE}

long_scores <- gather_scores(all_scores, "ngdp_rpch")
scores_to_table(all_scores, "ngdp_rpch")

@@

```{r gdptileplot, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

@@

```{r overall_assessment, fig.dim=c(12,12), echo = FALSE}

scores_cvgshort <- fread(here("quantile_forecasts", "ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE)

point_scores <- data.table::fread(here("quantile_forecasts", "pointfc_scores.csv"))

ovr_assessment_plot <- function(wis_scoredat, cvg_scoredat, point_scores, crps_scores, trgt){

  cvgscores <- cvg_scoredat |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  decomp_data <- wis_scoredat |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  point_scores <- point_scores |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, meanae := mean(ae), by = c("source", "horizon")) |>
    .d(, meansque := mean(sque), by = c("source", "horizon")) |>
    .d(, .(source, country, horizon, meanae, meansque)) |>
    unique()

  crps_scores_rw <- crps_scores |>
    .d(method == "rolling window") |>
    .d(target == trgt) |>
    .d(, target := NULL)

  crps_scores_loo <- crps_scores |>
    .d(method == "leave-one-out") |>
    .d(target == trgt) |>
    .d(, target := NULL)

  xmax <- max(decomp_data$interval_score)

  decomp_data <- decomp_data |>
    data.table::melt(id.vars = c("source", "error_method", "horizon"),
                     value.name = "score",
                     variable.name = "type") |>
    .d(type != "interval_score") # filter out total score, only interested in decomposition

  #check arguments of function!
  wis_plot <- function(decompdat, error_m, maxval){

    if(error_m == "directional"){
      plot_name <- "Directional Errors"
    } else {
      plot_name <- "Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "ar"),
                       labels = c("IMF", "BVAR", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_brewer(palette = "Set2", 
                        name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

    return(plt)

  }

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +
      scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none")

    if(!is.null(facet)){

      path_plot <- path_plot +
      facet_wrap(~ get(facet),
                 scales = "free") 
    }

    return(path_plot)

  }

  dir_wis_plot <- wis_plot(decomp_data, "directional", xmax)
  abs_wis_plot <- wis_plot(decomp_data, "absolute", xmax)

  c50plot <- horizon_path_plot(cvgscores, "coverage_50", "CI Coverage - 50", facet = "error_method") +
    geom_hline(aes(yintercept = 0.5)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  c80plot <- horizon_path_plot(cvgscores, "coverage_80", "CI Coverage - 80", facet = "error_method") +
    geom_hline(aes(yintercept = 0.8)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  erow <-  ggplot() + theme_void()

  aeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meanae", "Absolute Error")

  squeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meansque", "Squared Error")

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample")

  crpsplot_loo <- horizon_path_plot(crps_scores_loo |>.d(, error_method := "directional"), "score", "CRPS by Sample")

  ovr_plot <- 
    (dir_wis_plot | abs_wis_plot) /
    (plot_spacer()) /
    (plot_annotation(title = "hey")) /
    (c50plot | c80plot) /
    (plot_spacer()) /
    (squeplot | aeplot | crpsplot_rw | crpsplot_loo) +
    plot_layout(guides = "collect", 
                heights = c(0.2, 0.05, 0.2, 0.05, 0.2)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

  return(ovr_plot)
}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "pcpi_pch")

### 2023-09-12 manuscript/manuscript.tex
manuscript daily update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={IMF and Benchmark Forecasts},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

@@

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}

@@

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={IMF and Benchmark Forecasts},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#bring into long format}
\NormalTok{long\_scores }\OtherTok{\textless{}{-}} \FunctionTok{gather\_scores}\NormalTok{(all\_scores, }\StringTok{"pcpi\_pch"}\NormalTok{)}

\FunctionTok{scores\_to\_table}\NormalTok{(all\_scores, }\StringTok{"pcpi\_pch"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{manuscript_files/figure-latex/cpitileplot-1.pdf}

@@

\includegraphics{manuscript_files/figure-latex/gdptileplot-1.pdf}

\begin{verbatim}
## Warning: Using alpha for a discrete variable is not advised.
## Using alpha for a discrete variable is not advised.
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/overall_assessment-1.pdf}
